---
title: "Lab 05 - Grading the professor"
subtitle: "Modelling with R: regression"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: ../lab.css
link-citations: yes
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Many college courses conclude by giving students the opportunity to evaluate the course and the instructor anonymously. However, the use of these student evaluations as an indicator of course quality and teaching effectiveness is often criticized because these measures may reflect the influence of non-teaching related characteristics, such as the physical appearance of the instructor.

The article titled, "Beauty in the classroom: instructors' pulchritude and putative pedagogical productivity" (Hamermesh and Parker, 2005) found that instructors who are viewed to be better looking receive higher instructional ratings. (Daniel S. Hamermesh, Amy Parker, Beauty in the classroom: instructors pulchritude and putative pedagogical productivity, Economics of Education Review, Volume 24, Issue 4, August 2005, Pages 369-376, ISSN 0272-7757, 10.1016/j.econedurev.2004.07.013. <http://www.sciencedirect.com/science/article/pii/S0272775704001165>.)

In this lab you will analyze the data from this study in order to learn what goes into a positive professor evaluation. The data were gathered from end of semester student evaluations for a large sample of professors from the University of Texas at Austin. In addition, six students rated the professors' physical appearance. (This is a slightly modified version of the original data set that was released as part of the replication data for Data Analysis Using Regression and Multilevel/Hierarchical Models (Gelman and Hill, 2007).) The result is a data frame where each row contains a different course and columns represent variables about the courses and professors.

# Learning goals

-   Fitting a linear regression with a single numerical and categorical predictor, and with multiple predictors
-   Interpreting regression output in context of the data
-   Comparing models
-   Fitting a linear regression with multiple predictors


# Getting started

Go to the course GitHub organization and locate your homework repo, clone it in RStudio and open the R Markdown document.
Knit the document to make sure it compiles without errors.

## Warm up

- [Click here to go the the Lab in Canvas](https://georgetown.instructure.com/courses/126656/assignments/572748)
- Click on the link inside the assignment to go to the acceptance page. You should see something like the following screen, but the name of the repository that will be created is `lab-5-YOURGITHUBUSERNAME`:
- Grab the URL of the repo, and clone it in RStudio.
- Open the R Markdown document `lab-05.Rmd` and Knit it.
- Make sure it compiles without errors.
- The output will be in the file markdown `.md` file with the same name.

## Packages

```{r}
library(tidyverse) 
library(tidymodels)
library(openintro)
```


## Data

The data can be found in the **openintro** package, and it's called `evals`.
More information on the data is [here](https://www.openintro.org/data/index.php?data=evals).

#  Part I Exercises

## Exploratory Data Analysis

1.  Visualize the distribution of `score`.
    Is the distribution skewed?
    What does that tell you about how students rate courses?
    Is this what you expected to see?
    Why, or why not?
    Include any summary statistics and visualizations you use in your response.

2.  Visualize and describe the relationship between `score` and `bty_avg`.

```{marginfigure}
**Hint:** See the help page for the function at http://ggplot2.tidyverse.org/reference/index.html.
```

3.  Recreate the scatterplot from Exercise 2, but this time use\
    `geom_jitter()`? What does "jitter" mean? What was misleading about the initial scatterplot?

🧶 ✅ ⬆️ *If you haven't done so recently, knit, commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.*

## Linear regression with a numerical predictor

```{marginfigure}
Linear model is in the form $\hat{y} = b_0 + b_1 x$.
```

4.  Let's see if the apparent trend in the plot is something more than natural variation.
    Fit a linear model called `score_bty_fit` to predict average professor evaluation `score` by average beauty rating (`bty_avg`).
    Based on the regression output, write the linear model.

5.  Recreate the scatterplot from Exercise 2, and add the regression line to this plot in orange colour, with shading for the uncertainty of the line turned off.

6.  Interpret the slope of the linear model in context of the data.

7.  Interpret the intercept of the linear model in context of the data.
    Comment on whether or not the intercept makes sense in this context.

8.  Determine the $R^2$ of the model and interpret it in context of the data.

🧶 ✅ ⬆️ *If you haven't done so recently, knit, commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.*

## Linear regression with a categorical predictor

9.  Fit a new linear model called `score_bty_gender` to predict average professor evaluation `score` based on `gender` of the professor.
    Based on the regression output, write the linear model and interpret the slope and intercept in context of the data.

10. What is the equation of the line corresponding to male professors?
    What is it for female professors?

11. Fit a new linear model called `score_bty_rank` to predict average professor evaluation `score` based on `rank` of the professor.
    Based on the regression output, write the linear model and interpret the slopes and intercept in context of the data.

12. Create a new variable called `rank_relevel` where `"tenure track"` is the baseline level.

13. Fit a new linear model called `score_bty_rank_relevel` to predict average professor evaluation `score` based on `rank_relevel` of the professor.
    This is the new (releveled) variable you created in Exercise 12.
    Based on the regression output, write the linear model and interpret the slopes and intercept in context of the data.
    Also determine and interpret the $R^2$ of the model.

14. Create another new variable called `tenure_eligible` that labels `"teaching"` faculty as `"no"` and labels `"tenure track"` and `"tenured"` faculty as `"yes"`.

15. Fit a new linear model called `score_bty_tenure_eligible` to predict average professor evaluation `score` based on `tenure_eligible`ness of the professor.
    This is the new (regrouped) variable you created in the previous exercise.
    Based on the regression output, write the linear model and interpret the slopes and intercept in context of the data.
    Also determine and interpret the $R^2$ of the model.

🧶 ✅ ⬆️ Knit, *commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards and review the md document on GitHub to make sure you're happy with the final state of your work.*

# Part II Exercises

## Simple linear regression

2.  Fit a linear model (one you have fit before): `score_bty_fit`, predicting average professor evaluation `score` based on average beauty rating (`bty_avg`) only. Write the linear model, and note the $R^2$ and the adjusted $R^2$.

🧶 ✅ ⬆️ *If you haven't done so recently, knit, commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.*

## Multiple linear regression

3.  Fit a linear model (one you have fit before): `score_bty_gender_fit`, predicting average professor evaluation `score` based on average beauty rating (`bty_avg`) and `gender`.
    Write the linear model, and note the $R^2$ and the adjusted $R^2$.

4.  Interpret the slopes and intercept of `score_bty_gen_fit` in context of the data.

5.  What percent of the variability in `score` is explained by the model `score_bty_gen_fit`.

6.  What is the equation of the line corresponding to *just* male professors?

7.  For two professors who received the same beauty rating, which gender tends to have the higher course evaluation score?

8.  How does the relationship between beauty and evaluation score vary between male and female professors?

9.  How do the adjusted $R^2$ values of `score_bty_gen_fit` and `score_bty_fit` compare?
    What does this tell us about how useful `gender` is in explaining the variability in evaluation scores when we already have information on the beuaty score of the professor.

10. Compare the slopes of `bty_avg` under the two models (`score_bty_fit` and `score_bty_gen_fit`).
    Has the addition of `gender` to the model changed the parameter estimate (slope) for `bty_avg`?

11. Create a new model called `score_bty_rank_fit` with `gender` removed and `rank` added in.
    Write the equation of the linear model and interpret the slopes and intercept in context of the data.

🧶 ✅ ⬆️ Knit, *commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards and review the md document on GitHub to make sure you're happy with the final state of your work.*


